Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
fix up to false findings

**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.

**Residuals** summarise what is **not captured by the regression model**.

@@

.center[
.caption[
Fig. 1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
]
]

@@

.float-left[
But there are **drawbacks** of using them.

@@

.pull-left-center-80[

@@

```{r}
set.seed(10086)
# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)
# Base data
dat <- mod$gen(300)
bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
  facet_wrap(~type, ncol = 2, scales = "free") +
  theme(axis.text = element_blank()) +
  xlab("Fitted values") +
  ylab("Residuals")
```

@@

.pull-right-120[

BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).

@@

.center[
.caption[
Tab. 1: _Statistical significance testing for departures from good residuals. Shown are the \\(p\\)-values calculated for the RESET, the BP and the SW tests._
]
]

@@

@@

.center[
.caption[
Fig. 3: _Residual plot of a correctly specified model exhibiting a heteroskedasticity pattern (triangle shape)._
]
]

@@

# 🤔False Visual Discoveries🤔

@@

Can you find the evidence on heteroskedasticity?

>>>>

**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model. **Residuals** summarise what is **not captured by the regression model**.

@@

@@

.float-left[

They are **very good at detecting violations of model assumptions**, but there are some **unexpected usages**.

@@

@@

```{r}
set.seed(10086)
# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)
# Base data
dat <- mod$gen(300)
bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
  facet_wrap(~type, ncol = 2, scales = "free") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  xlab("Fitted values") +
  ylab("Residuals")
```

@@

All the hypothesis tests **correctly reject** the correspoding model defects, and **correctly not reject** the structure in residual plot (A).

BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for**.

@@

@@

.center[
# 🤔False Visual Discoveries🤔
]

@@

@@

@@

Can you find the **evidence** on heteroskedasticity?
